\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{Lecture 1}
  The goal of this class is to solve \emph{mathematical} problems with the help of \emph{computers}.
  \subsection{Chapter Summaries}
  \begin{enumerate}
    \item Computations in computers
    \begin{itemize}
      \item How to store (real) numbers in the computer
      \begin{note}{}
        If the number has finitely many digits, then it is simple. What about numbers with infinitely many digits, i.e. $\frac{1}{3}$? We have to truncate or round, and store an approximation with finitely many digits.
      \end{note}
      \item How to perform computations
      \begin{note}{}
        From regular math, we know that $\frac{1}{3} + \frac{1}{3} = \frac{2}{3}$. However, in the computer, due to errors, we have $\frac{1}{3}\oplus \frac{1}{3} = ? \oplus ? = ??$.
      \end{note}
      \item Errors
    \end{itemize}
    \item 
    \begin{itemize}
      \item Find roots of $f(x) = 0$ using bisection, Newton's method, ...
      \item Convergence
      \item Convergence order (how fast it converges)
    \end{itemize}
    \item Polynomial interpolation
    \begin{itemize}
      \item Approximate a function $f(x)$ by a polynomial $P(x)$, where $f(x_i) = P(x_i)$ for finitely many $x_i$
      \item \emph{Accuracy} of the polynomial approximations
    \end{itemize}
    \item Numerical differentiations and numerical integrations
    \begin{itemize}
      \item Using the approximations from chapter $3$, we can approximate using
      \[
        f'(x^*)\approx P'(x^*) = \sum_{i=0}^{k}f(x_k)c_k,
      \]
      and
      \[
        \int_{a}^{b}f(x) \,\mathrm dx\approx \int_{a}^{b}P(x) \,\mathrm dx = \sum_{i=1}^{k} f(\overline{x_k})\overline{c_k}.
      \]
      \item Error analysis
    \end{itemize}
    \item[6.7.] Solving linear systems of equations
    \begin{itemize}
      \item Direct methods: Gaussian elimination (computationally expensive)
      \item Iterative methods: (faster and cheaper)
      \item Solution stability
    \end{itemize}
  \end{enumerate}
  \subsection{Round-off errors and computer arithmetics}
  There are three kinds of errors:
  \begin{itemize}
    \item Modeling Error: Occurs when we convert a problem from the real world into the mathematical world.
    \item Method Error: Occurs when we try to solve the mathematical problem numerically.
    \item Round-off error: Occurs when the computer gives an incorrect result with the correct algorithm (comes from storage and computation).
  \end{itemize}
  \subsubsection{Storage}
  Infinite digit real numbers are \emph{stored} as finite digit numbers, using the normalized decimal form of real numbers.
  \begin{definition}{Normalized decimal form of a real number}
    For any $y\in\R$, we may write
    \[
      y = \pm 0.d_1d_2d_3\dotsc d_kd_{k + 1}\dotsc\cdot 10^n,
    \]
    where $0 < d_1\leq 9$, $0\leq d_i\leq 9$, $n$ are integers. For the particular case where $y = 0$, we write $y = 0.0\cdot 10^0$.
  \end{definition}
  \begin{definition}{Normalized machine numbers (Floating-point form)}
    Any machine number $y$ can be written as
    \[
      y = \pm 0.d_1d_2\dotsb d_k\cdot 10^n,
    \]
    where $0 < d_1\leq 9$, $0\leq d_i\leq 9$, $n$ are integers.
  \end{definition}
  We can think of the storage process as mapping normalized real numbers to normalized machine numbers. We do this via rounding or truncating. \par
  Consider some $y\in\R\sm \set{0}$.
  \begin{itemize}
    \item Truncating ($k$-digit truncation of $y = \pm 0.d_1d_2d_3\dotsc d_kd_{k + 1}d_{k + 2}\dotsc\cdot 10^n$)
    Simply omit the digits from $d_{k + 1}$ and onwards, in other words
    \[
      f_\ell(\pm 0.d_1d_2d_3\dotsc d_kd_{k + 1}d_{k + 2}\dotsc\cdot 10^n) = \pm 0.d_1d_2\dotsc d_k\cdot 10^n.
    \]
    Thus we have $y\approx f_\ell(y)$.
    \item Rounding ($k$-digit rounding of $y = \pm 0.d_1d_2d_3\dotsc d_kd_{k + 1}d_{k + 2}\dotsc\cdot 10^n$) \par
    If $d_{k + 1} < 5$, then we drop $d_{k + 1}d_{k + 2}\dotsc$ (same with truncating) \par
    If $d_{k + 1} \geq 5$, then add $1$ to $d_k$ and drop $d_{k + 1}d_{k + 2}\dotsc$
    \[
      f_\ell(\pm 0.d_1d_2d_3\dotsc d_kd_{k + 1}d_{k + 2}\dotsc\cdot 10^n) = \pm \delta_1\delta_2\dotsc \delta_k\cdot 10^m.
    \]
  \end{itemize}
\end{document}
