\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{Lecture 3}
  \textbf{Question.} Can we say more about what happens close to fixed points?
  \begin{example}{} Consider the equation given by
    \[
      \dot{x} = x(x + 1)(x - 1)^2,
    \]
    which has stable points at $-1$, $0$, and $1$. \par
    We will use something called the \emph{local method}. We define a new function $\eta(t) = x(t) + 1$, so $x(t) = \eta(t) - 1$. Hence $\dot{x}(t) = \dot{\eta}(t)$. Furthermore,
    \begin{align*}
      x(x + 1)(x - 1)^2 &= (\eta - 1)\eta(\eta - 2)^2 \\
                        &= -4\eta + O(\eta^2), \tag{$\eta\to 0$}
    \end{align*}
    so $\dot{\eta}\approx -4\eta$. Near $x = -1, \eta = x + 1$ and it satisfies $\dot{\eta} = -4\eta$, so $\eta(t)\approx Ce^{-4t}$. We can see that this approaches $0$ as $t\to\infty$, so points around $x = -1$ will approach $-1$.
  \end{example}
  In general, we have the following method: \par
  Assume that $x^*$ is a fixed point of $\dot{x} = f(x)$, i.e. $f(x^*) = 0$. Let $\eta = x - x^*$. Then
  \begin{align*}
    \dot{\eta} &= \dot{x} \\
            &= f(x) \\
            &= f(x^*) + f'(x^*)\eta + \frac{1}{2}f''(x^*)\eta^2 + \dotsb \\
            &= f'(x^*)\eta + O(\eta^2). \tag{$\eta\to 0$}
  \end{align*}
  Hence the equation
  \[
    \dot{\eta} = f'(x^*)\eta
  \]
  is the \emph{linearization} at $x = x^*$. We know that the solution to such a differential equation is
  \[
    \eta(t) = Ce^{f'(x^*)t} = \begin{cases}
      0, & f'(x^*) < 0 \\
      \pm\infty, & f'(x^*) > 0
    \end{cases},
  \]
  as $t\to\infty$. In the first case, the terms near $x^*$ will tend towards $x^*$, and in the latter they will diverge from $x^*$.
  \begin{theorem}{}
    Suppose that $x^*$ is a fixed point of the system $\dot{x} = f(x)$. Then if
    \begin{itemize}
      \item $f'(x^*) < 0$, the fixed point $x^*$ is stable.
      \item $f'(x^*) > 0$, the fixed point $x^*$ is unstable.
    \end{itemize}
  \end{theorem}
  \textbf{Question.} What happens if $f'(x^*) = 0$? Anything can happen/the test is inconclusive. \par
  \begin{itemize}
    \item Consider the equation $\dot{x} = x^3$. We see that $x^* = 0$ is a critical point, and that $f'(x) = 3x^2$, so $f'(x^*) = 0$. Using our usual graphical methods, we can see that points to the left of $x^*$ will approach $x^*$, and so will points to the right, and so $x^* = 0$ is a stable point.
    \item If we use the equation $\dot{x} = -x^3$, we get the direct opposite, that is $x^* = 0$ is unstable despite having the same critical point.
    \item If we look at the behavior of $\dot{x} = x^2$, then we get that the critical point at $x^* = 0$ is half-stable.
    \item If we consider the equation $\dot{x} = 0$, then every number on the real line is a critical point, and so the solutions don't move at all.
  \end{itemize}
  \subsection{Potentials}
  \begin{itemize}
    \item Let $f\colon \R\to \R$ be smooth and consider the system
    \[
      \dot{x} = f(x).
    \]
    \item A function $V\colon \R\to \R$ so that
    \[
      f(x) = -V'(x)
    \]
    is called a \emph{potential} for $f$.
    \item Our system can be written as a \emph{gradient flow}
    \[
      \dot{x} = -V'(x).
    \]
  \end{itemize}
  \begin{note}{}
    Potential functions are \emph{not} unique, since you can always add a constant.
  \end{note}
\end{document}
