\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{Lecture 28}
  \subsection{Lens-Maker's Equation}
  The Lens-maker's equation is given by
  \[
    \frac{1}{f} = (n-1) \paren{\frac{1}{R_1} + \frac{1}{R_2}},
  \]
  assuming that your lens is in air. We call $\frac{1}{f}$ the ``bending power'', whose units are metres\tsup{-1}, or diopters.
  \subsection{Cameras}
  The focal length of the camera lens should approximately be equal to the distance from the lens to the screen, to be able to focus images onto the screen.
  \begin{note}{}
    In reality the focal distance should be slightly less because the object is not always at infinity, and should be adjustable.
  \end{note}
  Pinhole cameras don't need to be focused, but you need a very small hole to produce a sharp image, but this reduces the amount of light that the camera can get.
  \begin{definition}{F-stop}
    The F-stop is defined by the distance to the back of the box over the diameter of the lens, or
    \[
      \text{f-stop} \ceq \frac{d}{d_\text{lens}}.
    \]
  \end{definition}
  \begin{note}{}
    A smaller f-stop is generally better because it implies a larger lens, but the small-angle approximation become worse and worse. Real cameras use multiple lenses to get a better image.
  \end{note}
  \subsection{Magnifying Lenses}
  The goal of a magnifying lens is not to make the image larger, but rather allow you to focus on something that is closer. The book derives that
  \[
    M \ceq \frac{\theta'}{\theta} = \frac{25\ \mathrm{cm}}{f}. \tag{object at $\infty$}
  \]
\end{document}
