\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{Lecture 4}
  \begin{itemize}
    \item There are $n^r$ \emph{ordered} samples of size $r$ from $n$ objects \emph{with replacement}.
    \item There are $\phantom{}_nP_r$ \emph{ordered} samples of size $r$ from $n$ objects \emph{without replacement}.
    \item There are $\phantom{}_nC_r = \frac{n!}{r!(n - r)!}$ \emph{unordered} samples of size $r$ from $n$ objects \emph{without replacement}.
    \item There are $\phantom{}_{n + r - 1}C_r = \frac{n!}{r!(n - r)!}$ \emph{unordered} samples of size $r$ from $n$ objects \emph{with replacement}.
  \end{itemize}
  \subsection{Distinguishable Permutations}
  \begin{itemize}
    \item Suppose we are given $n$ objects, but some of them are identical.
    \item How many \emph{distinguishable} permutations of the $n$ objects are there?
  \end{itemize}
  \begin{theorem}{}
    Suppose you have:
    \begin{itemize}
      \item $n_1$ objects of type $1$,
      \item $n_2$ objects of type $2$,
      \item ...
      \item $n_r$ objects of type $r$.
    \end{itemize}
    Let $n = n_1 + n_2 + \dotsb + n_r$. Then the number of distinguishable permutations is
    \[
      \binom{n}{n_1,n_2,\dotsc,n_r} = \frac{n!}{n_1!n_2!\dotsb n_r!}.
    \]
    \begin{proof}
      We have $n$ locations.
      \begin{itemize}
        \item First choose $n_1$ locations for type $1$ objects: $\choose{n}{n_1}$ choices.
        \item Then choose $n_2$ locations for type $2$ objects: $\choose{n - n_1}{n_2}$ choices.
        \item $\dotsb$
        \item Finally we choose $n_r$ locations for type $r$ objects: $\choose{n - n_1 - \dotsb - n_{r - 1}}{n_r}$ choices.
      \end{itemize}
      Using the multiplication principle to take the product of all of these combinations, we have
      \[
        \binom{n}{n_1,\dotsc,n_r} = \frac{n!}{n_1!n_2!\dotsb n_r!}.
      \]
    \end{proof}
  \end{theorem}
  \begin{note}{}
    An alternate way to think about this theorem is to first consider how many regular permutations of $n$ objects there are ($n!$), and then divide by how many possible times we over count ($n_k!$ for each $1\leq k\leq r$). 
  \end{note}
  \subsection{The Binomial Theorem}
  \begin{theorem}{Binomial Theorem}
    If $n\geq 0$ then
    \[
      (x + y)^n = \sum_{r=0}^{n}\binom{n}{r}x^ry^{n - r},
    \]
    where the \emph{binomial coefficient} is
    \[
      \binom{n}{r} = \choose{n}{r}.
    \]
    \begin{proof}
      If we multiply out $(x + y)^n = \underbrace{(x + y)\dotsb(x + y)}_{n\text{ times}}$ without using the fact that multiplication is commutative, we see that the number of times $x^ry^{n - r}$ appears is equal to how many different ways there are to rearrange $r$ ``$x$'' terms in $n$ total terms.
    \end{proof}
  \end{theorem}
  \begin{theorem}{}
    We have
    \[
      \sum_{r=0}^{n} \binom{n}{r} = 2^n.
    \]
    % \begin{proof}
    %   Observe that
    %   \[
    %     \sum_{r=0}^{n}\binom{n}{r} = \sum_{r=0}^{n}\binom{n}{r}1^r1^{n - r} = (1 + 1)^n = 2^n.
    %   \]
    % \end{proof}
  \end{theorem}
  \begin{theorem}{}
    If $n, r\geq 0$ then
    \[
      (x_1 + x_2 + \dotsb + x_r)^n = \sum_{n_1 + \dotsb + n_r = n}\binom{n}{n_1, n_2, \dotsc, n_r}x_1^{n_1}x_2^{n_2}\dotsb x_r^{n_r}.
    \]
    \begin{proof}
      Similar to the binomial theorem, we have that to get each term we just need to find the number of distinguishable permutations of $n_1$ terms of $x_1$, ..., $n_r$ terms of $x_r$, which is our multinomial coefficient from before.
    \end{proof}
  \end{theorem}
\end{document}
