\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{Lecture 16}
  Most ``rocky'' exoplanets are more massive than the Earth. We can look at emission spectra to see what elements are a part of the atmosphere, to detect potentially habitable planets.
  \subsection{Why Image Young Planets?}
  \begin{itemize}
    \item Much more radiation expected from hotter, younger planets
    \item Young planets are the only ones that we can image directly
  \end{itemize}
  Most normal stars are fusing hydrogen and helium in their cores (just like the Sun). The biggest red stars have a radius 1000x that of the Sun. The smallest red stars have a radius 0.1x that of the Sun.
  \subsection{Mass and Lifetime of Stars}
  If a star has 10x the mass of the Sun, then it has 10x as much fuel, but uses it 10\tsup{4} times as fast. If a star has 0.1x the mass of the Sun, then it has 0.1x the fuel, but uses it $0.001$ times as fast.
  \begin{note}{}
    The habitable zones of stars range with the mass of the star.
  \end{note}
  \subsection{Binary Stars}
  About 50\% of stars are formed in binary systems
  \begin{itemize}
    \item For it to host complex life, both stars must be low mass
    \item The system lifetime is limited by the evolution of the most massive star
    \item If both stars are on the main sequence then you're more likely to have a habitable planet
  \end{itemize}
  The \emph{Hertzprung-Russell Diagram} plots the luminosity of stars as functions of their temperature. The temperature decreases from left to right.
  \begin{note}{}
    Low-mass stars are more likely to be seen than high-mass stars.
  \end{note}
  Stellar properties depend on both mass and age---stars that have exhausted their supplies of hydrogen and helium are no longer on the main sequence. All stars becomes larger and redder after exhausting their core hydrogen. Low-mass stars on the main sequence will change the least over time.
  \begin{note}{}
    The life of a star is determined by its starting mass.
  \end{note}
  \subsection{Star Clusters}
  \begin{itemize}
    \item Open clusters are groups of a few thousand loosely packed stars
    \item Globular clusters are tightly packed groups of around half a million stars
    \item The stars in each cluster usually form around the same time
    \item More blue stars implies younger cluster
    \begin{note}{}
      Thus the cluster is the same age as the age of the hottest star.
    \end{note}
  \end{itemize}
  Degeneracy pressure does not depend on the temperature.
  \subsection{Brown Dwarfs}
  \begin{itemize}
    \item An object less massive than $0.08$ M\tsub{sun}
    \item Radiates mostly infrared light
    \item No heat from fusion
    \item Brown dwarfs cool off after degeneracy pressure stops gravitational contraction
  \end{itemize}
  \subsection{3-alpha process}
  Two Helium collide to form a Beryllium atom, which then collides with a third helium atom to make a carbon atom. This requires very high temperatures to occur (100M Kelvin). \par
  Hydrogen fuses in a shell around the Helium core once it's too cold to continue the 3-alpha process. \par
  When the core temperature rises enough to fuse helium again, helium fusion rises very sharply (helium flash). \par
  Helium burning stars neither shrink nor grow because the thermostat is temporarily fixed. Helium fuses into carbon into the core. When the core becomes carbon, the star stars fusing helium in a shell around the carbon core. These double shells generate enough heat to push away the outer layers of the star, which results in a \emph{planetary nebula}. The small hot core is left, which is a white dwarf.
  \begin{note}{}
    When the Sun uses up the nuclear fuel, it will expand and become larger. It will become brighter.
  \end{note}
\end{document}
