\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{Lecture 18}
  The LIGO observatory measures wobbles of space-time from merging black holes. We can use LIGO to observe these gravitational waves traverse space-time at the speed of light.
  \subsection{Where are the Aliens?}
  Fermi's paradox: Plausible arguments suggest that aliens should be common. So why haven't we detected them? \par
  No substantiated evidence for visitations has been presented
  \subsection{Interstellar Travel}
  Rockets achieve thrust by igniting mixtures of chemical propellants and expelling them at high velocity (Newton's 3\tsup{rd} law). Rockets are limited by the mass ratio (ratio of the mass of the rocket before and after its fuel is burned). Reaching escape velocity requires a mass ratio of roughly $39$, so single stage rockets need to carry 38x as much fuel as rocket. \par
  With conventional technology, higher energy efficiency can be achieved with nuclear fission. Nuclear fission has an efficiency of roughly 0.07\%, whereas fusion has an efficiency of roughly 0.7\%.
  \begin{itemize}
    \item Ion drives are low-power, high-efficiency options
    \item Solar sails catch the solar wind for thrust (perhaps with a terrestrial laser)
    \item Interstellar arks might put people into \emph{suspended hibernation}
    \item Generation ships have multiple generations of people live and die on the voyage to a new home
  \end{itemize}
  \subsubsection{Relativity in Interstellar Travel}
  \begin{itemize}
    \item Time dilation is a natural consequence of moving close to the speed of light
  \end{itemize}
  Energy concerns---nuclear fusion would be limited in achieving relativistic speeds. Interstellar ramjets could work, as well as using matter-antimatter annihilations. \par
  Colonisation is part of humanity's drive for exploration.
  \subsection{Solutions to the Fermi Paradox}
  \begin{itemize}
    \item We are alone (rare Earth hypothesis)
    \item Civilisations, but no colonisation---perhaps technological difficulties or sociological difficulties, or there is a galactic civilisation that is indifferent (or treats us like zoo animals or only appears once we reach a technological level)
  \end{itemize}
  \begin{theorem}{Rocket Equation}
    For initial and final masses $M_i$ and $M_f$, and exhaust speed $v_c$, we have
    \[
      v = v_c\ln \frac{M_i}{M_f}.
    \]
  \end{theorem}
  Perhaps civilisations are common but interstellar travel is not. \par
  Mars, Europa, Enceladus are at the top of the list for habitability.
\end{document}
