\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{Lecture 1}
  \subsection{Overview of This Course}
  ``Ancient'' history of mathematics
  \begin{itemize}
    \item Egyptian mathematics (c. $3200$ BCE -- $300$ BCE): Number system in base $10$, fractions.
    \item Babylonian mathematics (c. $3300$ BCE -- $500$ BCE): Place-value sexagesimal number system, roots of algebra.
    \item Greek mathematics (c. $700$ BCE -- $400$ CE): Deduction method, geometry, roots of number theory.
    \item Mathematics in Asia (c. $1100$ BCE -- $1200$ CE): Roots of number theory and algebra.
    \item Polynomial equations.
    \item Development of calculus, infinite series (16\tsup{th}--17\tsup{th} centuries CE).
  \end{itemize}
  \subsection{The Beginnings}
  What is ``Mathematics''?
  \begin{enumerate}
    \item Logical-deductive mathematics (initiated by the Greeks). This is not the generally accepted view of ``mathematics'' anymore.
    \item Abstract counting problems: ``mathematical'' exercises without a direct practical use.
    \item Practical counting problems: Bookkeeping for bureaucracy, inventories of goods and harvest, calculations involving lengths and areas for farming purposes, etc.
    \item Development of a number system.
  \end{enumerate}
  \begin{note}{}
    We will see later that 2--4 all occur almost simultaneously.
  \end{note}
  \subsection{Egyptian Mathematics}
  Most writing was done on papyrus, which doesn't preserve well, so we don't have much evidence from this period.
  \subsubsection{Egyptian Number System}
  The Egyptians had a base $10$ number system, with a new Hieroglyphic symbol for each power of $10$. This number system is \emph{not} place valued. One simply adds all the symbols to get the number. For instance $\cap\cap\cap|||| = 34$. By comparison, our modern number system is place-valued, which means that the position of each digit matters, i.e. $254 \neq 452$.
  \subsubsection{Egyptian Arithmetic}
  Summation is extremely easy, you just write the numbers next to each other to get their sum. Multiplication was done via consecutive doubling and adding. We do this by decomposing one number into powers of $2$, and then distributing the other number across this sum. Because of this method, Egyptians had tablets with powers of $2$, and tablets with consecutive doublings of many numbers. \par
  We also have sources of:
  \begin{itemize}
    \item Fractions---The Egyptians used almost exclusively fractions of the form $\frac{1}{n}$ (aside from $\frac{2}{3}$ and $\frac{3}{4}$) and used sums of these to express more general fractions.
    \item Notation---Since any fraction can be written as the sum of fractions of the form $\frac{1}{n}$, any general fraction could be written by just writing $\frac{1}{n}$ fractions next to each other.
    \item Algebra---We have evidence that they solved some linear and second order (quadratic) equations.
    \item Geometry---Though limited, we have evidence that the Egyptians could calculate areas and volumes of triangles, circles, pyramids, etc.
  \end{itemize}
  However we see no evidence for a general theory for solving these questions, and all of the problems that we know of are elementary and mostly practical. The earliest ``advanced'' mathematical resources from the Egyptians are from around $1900$ BCE, and are predated by Babylonian mathematics.
\end{document}
