\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{Lecture 19}
  \begin{note}{}
    If $k > 1$, we can sometimes use it to find solutions.
  \end{note}
  Suppose $x_1^2 - ny_1^2 = k$. Composing with itself, we get a solution $(x_3, y_3)$ to $x^2 - ny^2 = k^2$, i.e.
  \[
    \frac{x_3^2}{k^2} - \frac{ny_3^2}{k^2} = 1,
  \]
  so $\paren{\frac{x_3}{k}, \frac{y_3}{k}}$ is a solution to the Pell equation.
  \begin{note}{}
    This is a \emph{rational} solution to the Pell equation, but not necessarily an \emph{integer} solution.
  \end{note}
  We hope that by composing a rational solution with itself enough times, we can find an integer solution.
  \begin{example}{}
    Consider the equation
    \[
      x^2 - 92y^2 = 1.
    \]
    We take our first guess to be ``small'', so $(x_1, y_1) = (10, 1)$. Hence we have $x_1^2 - 92y_1^2 = 8$. Composing this solution with itself, we have $(10\cdot 10 + 92\cdot 1\cdot 1, 10\cdot 1 + 1\cdot 10, 8^2) = (192, 20, 64)$. Dividing by $8^2$, we get the rational solution $(24, \frac{5}{2}, 1)$. We can then compose this new solution with itself, yielding $(24\cdot 24 + 92\cdot \frac{5}{2}\cdot \frac{5}{2}, 24\cdot \frac{5}{2} + \frac{5}{2}\cdot 24, 1^2) = (1151, 460, 1)$. Therefore we have $(x, y, k) = (1151, 120, 1)$ is an integer solution to
    \[
      x^2 - 92y^2 = 1.
    \]
  \end{example}
  \textbf{Question.} Does this method always work? \par
  No, but it does if we can find $x, y$ such that $x^2 - ny^2 = k$ for $k = \pm 1, \pm 2, \pm 4$. \par
  500 years later, Bhaskara II discovered a new method, called the ``cyclic process'', which always leads to a solution where $k = \pm 1, \pm 2, \pm 4$.
  \begin{note}{}
    A proof that this always works was only given in the 18\tsup{th} century by Lagrange.
  \end{note}
  \subsection{Bhaskara's Method}
  \begin{itemize}
    \item Choose $a, b$ that are co prime, so $\gcd(a, b) = 1$, and $k \geq 1$ such that $a^2 - nb^2 = k$. In other words, the triple $(a, b, k)$ is a solution to the equation.
    \item Choose $m$ and consider the triple $(m, 1, m^2 - n)$.
    \item We can then compose them to get the triple $(am + nb, a + bm, k(m^2 - n))$.
    \item Divide all sides of the equation by $k^2$ to get
    \[
      \paren{\frac{am + nb}{k}}^2 - n\paren{\frac{a + bm}{k}}^2 = \frac{m^2 - n}{k}.
    \]
    If we choose $m$ such that $\frac{m^2 - n}{k}$ is an integer, then also $\frac{am + nb}{k}$ and $\frac{a + bm}{k}$ are integers as well.
  \end{itemize}
  \textbf{Idea.} Choose $m$ in such a way that $\frac{m^2 - n}{k}$ is as small as possible. \\
  \textbf{Fact.} Repeating this process eventually yields a solution with $\frac{m^2 - n}{k} = \pm 1, \pm 2, \pm 4$. 
\end{document}
