\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{Lecture 2}
  \subsection{Babylonian Mathematics}
  The Babylonians, in comparison to the Egyptians:
  \begin{itemize}
    \item had a more advanced number system,
    \item solved more difficult problems,
    \item solved more abstract (``useless'') problems.
  \end{itemize}
  Hence people often view the Babylonians in Mesopotamia as being the starting point of ``serious'' mathematics. \par
  Mesopotamia was located between the Tigris and Euphrates rivers, and its name means ``the land between the rivers''. Historical texts from Mesopotamia are unusually well-preserved, because texts were made by making impressions on clay tablets, which hardened and preserve well.
  \subsubsection{Babylonian Number System}
  The Babylonian number system was a \emph{sexagesimal} (base 60), place-valued system.
  \begin{note}{}
    \begin{itemize}
      \item The number zero didn't exist just yet! So instead of writing ``20'', they would just write ``2'', and figure out what number it took from context.
      \item They also had fractions, but didn't have a symbol for a decimal point, so they again guessed based on context.
      \item Furthermore, since there is no zero, there was no way of distinguishing between $1\cdot 60^2 + 30\cdot 1$, or $1\cdot 60 + 30\cdot \frac{1}{60}$, etc.
    \end{itemize}
  \end{note}
  Later on in Mesopotamia, we see the first usage of the number zero (300 BCE), but only as a place holder to solve the aforementioned issues. This was also done by the Mayans around $400$ CE. It was not until $600$ CE that the number zero was invented in India, and carried its full weight as a ``number''.
  \begin{note}{}
    There are still remnants of the Babylonian sexagesimal number system in the modern day.
    \begin{itemize}
      \item Time: Hours and minutes are divided into $60$ pieces
      \item Angle measurements are still done in multiples/factors of $60$, i.e. $360^\circ$.
    \end{itemize}
  \end{note}
  \subsubsection{Babylonian Algebra}
  This notation was also used to solve basic algebra questions, but we don't know how exactly these problems were solved (probably the same method as the Egyptians, guess and check). It is believed that math problems were given to students in order to train them in literacy, numeracy, and applications to administration. The Babylonians also knew about the quadratic formula, although it was described in words as opposed to being given as an equation. Furthermore, although they knew of the quadratic formula, they did not have negative numbers, and so found only one solution to quadratic equations. This is the first instance we know of where a civilization has an ``algorithmic'' solution to a problem.
  \subsubsection{Babylonian Geometry}
  We have some sources of Babylonian geometry that indicate that they knew of the Pythagorean Theorem. One example of this is the ``IM 67118'' tablet, which gives the sides of a rectangle given its area and diagonal length. There is a tablet called ``Plimpton 322'' that contains Pythagorean triples, or integer triplets $a, b, c$ satisfying $a^2 + b^2 = c^2$.
\end{document}
