\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{Lecture 25}
  \subsection{Differentiation}
  When differentiation was introduced in the early 17\tsup{th} century, the notion of a limit did not exist yet. Fermat introduced the idea of adding an $E$ as an ``infinitesimal'', and then just ignoring it later. Even though it wasn't rigorous at the time, it was made much more rigorous later on.
  \begin{note}{}
    This method only really works for polynomials. It can also be extended to more variables, which is basically a form of implicit differentiation.
  \end{note}
  \subsubsection{John Wallis 1616--1703}
  He was one of the first to try to completely ``arithmetize'' the theory of areas and volumes.
  \begin{itemize}
    \item He proved that $\int_{0}^{1}x^p \,\mathrm dx = \frac{1}{p + 1}$.
    \item He considered a new approach towards fractional powers, by considering $x^{\frac{1}{n}}$ as complementary to $x^n$.
    \item He claimed a lot of results ``by induction'', but didn't show that much work.
  \end{itemize}
  \subsection{Newton's Calculus}
  Calculus is an algebra of infinite series.
  \begin{itemize}
    \item The main problem is to represent functions as infinite series (of polynomials).
    \item The starting point for a lot of Newton's calculus was the binomial theorem for $(1 + x)^n$, which he extended to fractional values of $n$.
    \item He learned of Fermat's techniques for polynomial differentiation, which he considered to be basically complete, and so turned most of his efforts to integration.
    \item He introduced the chain rule.
    \item Motivated by infinite series, he was able to invert infinite series to find equations for other functions (i.e. using the series for $\ln x$ to compute a series for $e^x$).
    \item He also discovered the \emph{fundamental theorem of calculus}, unifying the theories of differentiation and integration.
  \end{itemize}
  \subsection{Leibniz's Calculus}
  Leibniz developed his calculus independently from Newton. Here are some main differences:
  \begin{itemize}
    \item Leibniz had a different (and better) notation.
    \item His work was more on the level of ``ideas'', rather than detailed proofs or computations. It lacks much of the depth and rigor of Newton's works.
    \item Leibniz preferred ``closed forms'' over infinite series representations.
  \end{itemize}
  Modern calculus uses Leibniz's ideas and notation, but Newton's computations. He introduced the following ideas:
  \begin{itemize}
    \item He was the first to use the word ``function'', and to think about calculus in terms of functions.
    \item Introduced the notation $\frac{\mathrm{d}y}{\mathrm{d}x}$, and actually thought of it as a real quotient of infinitesimals.
    \item Introduced the integral sign $\int$, and thought of it as an elongated ``s'' for ``sum''.
    \begin{note}{}
      The notation 
      \[
        \int f(x)\,\mathrm{d}x
      \]
      actually means to multiply the function ``height'' by an infinitesimal width.
    \end{note}
    \item He derived the fundamental theorem of calculus by observing that sums and differences cancel each other out.
  \end{itemize}
  Newton used $\dot{x}$ for derivatives, while Leibniz introduced $f'$ notation.
\end{document}
