\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{The Occupation}
  \subsection{The End of the War}
  \begin{itemize}
    \item Daily air raids by US bombers from Nov. $1944$ to Aug. 1945
    \item Bombing of Tokyo (Mar. 9--10, 1945)
    \begin{itemize}
      \item Over 100K dead, 1M homeless, $16$ square miles destroyed
    \end{itemize}
    \item Battle of Okinawa (March--July 1945)
    \begin{itemize}
      \item Over 250K dead (including 150K civilians)
    \end{itemize}
    \item Atomic bombs on August 6\tsup{th} and 9\tsup{th} $1945$ at Hiroshima and Nagasaki, respectively
    \item Japan surrenders on August $15$, 1945
    \item Around 40\% of Japan's urban area is destroyed, with nine million homeless
  \end{itemize}
  \subsection{The Occupation}
  \begin{note}{}
    The goal was to demilitarise Japan and keep Japan from its imperialistic tendencies. It was also supposed to make Japan more democratic.
  \end{note}
  \begin{itemize}
    \item Establishment of the UN in April of 1945
    \item Supreme Commander for the Allied Powers is General Douglas MacArthur
    \item Military Tribunals
    \item $1947$ Constitution
    \begin{itemize}
      \item Emperor symbol of the state
      \item Renunciation of War and armed forces
      \item Rights of people: universal suffrage
    \end{itemize}
    \item Japan--US Security Treaty (1951--1952)
  \end{itemize}
  \subsection{Immediate Aftermath of the War}
  \begin{itemize}
    \item 9M homeless, extreme poverty
    \item Average household spent 68\% of income on food in 1946
    \item Average height/weight of schoolchildren decreased until 1948
    \begin{note}{}
      Malnourishment was a major problem during this period.
    \end{note}
    \item Barter economy, black markets, paralysed rationing system
    \item Prostitution system for 300K occupying troops
  \end{itemize}
  \begin{note}{}
    Many ideas/groups in Japan ``as we know them'' arise during this time period.
  \end{note}
  \subsection{Postwar Government}
  \begin{itemize}
    \item Executive Branch
    \begin{itemize}
      \item Emperor, Prime minister (appointed by Emperor as directed by Diet), and Cabinet (Appointed by Prime minister)
    \end{itemize}
    \item Legislative Branch (Diet)
    \begin{itemize}
      \item House of Councilors (upper house)
      \item House of Representatives (lower house)
    \end{itemize}
    \item Judicial Branch
    \begin{itemize}
      \item Judges appointed by Emperor as directed by Diet
    \end{itemize}
  \end{itemize}
  New political parties are formed that are essentially descendants of the old political parties that were banned in 1942.
  \begin{note}{}
    Preamble of the $1947$ Constitution includes a clause saying that ``all nations deserve to be respected'', a stark contrast to the original Japanese ideals.
  \end{note}
  \subsection{Reforms During Occupation}
  \begin{itemize}
    \item Reintroduction of electoral politics/political parties
    \item Education system restructured to be more like US education, and militaristic/authoritarian content is purged
    \item Language reform (writing system simplified/standardised)
    \item Unsuccessful attempt to decentralise economy
    \begin{note}{}
      Different from the centralised war economy.
    \end{note}
    \item Initial support of labour movements, which are later repressed
  \end{itemize}
  During the Cold War, Japan becomes the US ``bulwark'' against communism in Asia. There is censorship and repression by occupation forces.
  \subsection{Postwar Economic Recovery}
  \begin{itemize}
    \item Korean War (1950--1953)
    \item Shipbuilding, iron and steel production, manufacturing (decline of agriculture)
    \item Annual growth of 10\% by $1955$ (to 1974)
    \item Auto mobile industry, electronics
    \item Life expectancy rises dramatically (from $47$ in $1935$ to $68$ in $1960$)
    \item Environmental problems
    \begin{itemize}
      \item Pollution
      \item Minamata disease
    \end{itemize}
  \end{itemize}
  \subsection{Postwar Society}
  \begin{itemize}
    \item Initial ``New Deal'' progressive reforms abandoned with Cold War
    \item Increased urbanisation
    \item Breaking down of traditional extended family
    \item Rights of women
    \item Cultural nationalism
    \begin{itemize}
      \item Japanese exceptionalism/uniqueness
    \end{itemize}
  \end{itemize}
\end{document}
