\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{Early Japan}
  \subsection{Geography of Japan}
  Japan sits on the intersection of four different tectonic plates, so it is in an earthquake-prone area (called the Ring of Fire). Japan used to be connected to the mainland, but got separated a while ago. Furthermore, the coast of Japan (that is now underwater) used to not be underwater.
  \subsection{The Paleolithic (pre-ceramic) Time Period}
  \begin{itemize}
    \item This time period was from $32000$ to $13000$ years ago (end of Pleistocene)
    \item Humans lived in rock shelters and caves.
    \item People were hunter-gatherers---they hunted large animals like Nauman's elephant, Giant deer, and Mammoths (in Hokkaido).
  \end{itemize}
  There aren't many remains left of this time period.
  \subsection{The J\=omon Period (12000 - 400 BCE)}
  \begin{itemize}
    \item J\=omon means ``rope pattern'' and refers to a style of clay pottery.
    \item People were still a part of hunter-gatherer societies.
  \end{itemize}
  \subsubsection{J\=omon Technology}
  \begin{itemize}
    \item Pottery had a technological purpose
    \begin{itemize}
      \item Food preservation, cooking
      \item Salt production (food preservation)
      \item Trade from the coast to Inland \\[10pt]
      Pottery also had cultural and symbolic meaning, by transmitting patterns from one generation to the next.
    \end{itemize}
    \item Hunting
    \begin{itemize}
      \item Traps, Hunting tools, Arrows
      \item Boar, deer, fish
    \end{itemize}
    \item The people began to settle in multiple households and built storage buildings
    \item The population grew from $22000$ to $360000$ from $6000$ BCE to $3500$ BCE
  \end{itemize}
  \subsection{The Yayoi Period (400 or earlier to 200 BCE)}
  \subsubsection{Key Facts}
  \begin{itemize}
    \item ``Yayoi'' is a neighbourhood of Tokyo
    \item Spread from Ky\=ush\=u north and east
    \item Migrants from Korean peninsula mixing with J\=omon?
    \item New technologies for farming and warfare
    \item The emergence of political units
  \end{itemize}
  \subsubsection{Yayoi Technology}
  \begin{itemize}
    \item Plant cultivation
    \begin{itemize}
      \item Dry cultivation (millet, barley, wheat, etc.)
      \item Wet cultivation (rice agriculture)
    \end{itemize}
    \item Use of iron, bronze, glass, cloth, and wood
  \end{itemize}
  \subsubsection{Yayoi Society}
  \begin{itemize}
    \item Increasingly hierarchical (some people are designated as ``special'')
    \item Evidence of warfare (weapons, fortifications, signs of violence in skeletal remains)
    \item Contact in the Korean peninsula, appearance in Chinese records
  \end{itemize}
  One characteristic of the Yayoi is that they made large bronze bell-shaped items that they buried underground. They could have been used to represent power. The Yayoi also buried their dead in a jar.
  \begin{note}{}
    Japan first appears in a text from the Wei Dynasty, where the Chinese document things about Japan.
  \end{note}
  \subsection{The Kofun Period (mid 200s to late 500s)}
  \begin{itemize}
    \item ``Kofun'' are large mounded tombs.
    \item The size of the mounds represents the status of the deceased.
    \item There's a lot of influence from Korean burial methods.
    \item The graves contained mirrors, swords, armour, and saddles.
    \item There were also clay statues called ``haniwa'', which were put on the ground surrounding the tomb.
  \end{itemize}
  \subsubsection{Kofun Culture}
  \begin{itemize}
    \item Frequent warfare (especially in the 300-400s).
    \item Frequent contact with Korean kingdoms and Chinese imperial states (400s onwards).
    \item Early stages of centralised state (late 400s, 500s).
    \item Emergence of Yamato court in the Kinai
    \begin{itemize}
      \item Conquest of other regions on Japanese islands.
      \item Very limited, highly specialised use of writing.
      \item Elite lineages called ``uji''(clans), formed by groups specialising in different occupations.
      \item Emergence of a hereditary royal lineage.
    \end{itemize}
  \end{itemize}
\end{document}
