\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{Lecture 13}
  \subsection{Ancient Chinese Mathematics}
  The ancient Chinese civilization centered around the ``Yellow River'', which is the 6\tsup{th} longest river in the world.
  \subsubsection{Numbers and Patterns}
  \begin{itemize}
    \item Compared to many other civilizations, the Chinese had a great interest in numbers and patterns.
    \item Numbers were believed to have cosmic significance. For instance, ``1'' is often related to water, and ``9'' to fire.
    \item Interesting patterns of numbers were also often given divine significance.
    \item One very famous example is the ``Lo Shu Magic Square'' (c. 650 BCE). Every row, column, and diagonal sums to the same number (15).
    \begin{itemize}
      \item There are several legends surrounding this square, usually involving a turtle (an animal regarded as symbolizing the world, the universe, immortality, wisdom, etc.)
      \item One such legend was that the river God kept flooding the Yellow River. It was due to a turtle with the magic square on its back that the emperor realized that he had to make 15 offerings to appease the river God.
      \item The magic square is unique, aside from mirroring and rotating.
    \end{itemize}
  \end{itemize}
  \subsubsection{Chinese Number System}
  Already around 2000 BCE, the Chinese had a decimal place-value system! \par
  They arranged bamboo rods to represent the numbers 1 to 9, and then placing them in columns, the different places corresponded to different powers of 10.
  \begin{note}{}
    There is still no ``0'', so there's no way to express numbers such as 102.
  \end{note}
  \subsubsection{Resources}
  \begin{itemize}
    \item The Chinese usually wrote on bamboo, so we have limited resources. Fortunately, we have \emph{some} resources dating back to 200 BCE. We have ``Jiuzhang Suanshu'', or ``Nine Chapters on the Mathematical Art'', written most likely over a long period of time between 1000 BCE to 250 AD by many different authors.
    \item Most of the problems discussed in this book are practical, revolving around solving real-world problems. They found \emph{general} methods for solving problems, rather than solutions to specific problems.
    \item The book covers:
    \begin{itemize}
      \item Arithmetic operations with positive integers, fractions, and some ``special'' irrational numbers. We even see negative numbers (usually in the context of debt)!
      \item It contains an approximation of $\pi \approx 3.14159$, using an approximation of the circle with a regular polygon of 192 vertices. We actually know the author for this part, Liu Hui (c. 250 AD).
      \item The Pythagorean Theorem.
      \item Solving linear systems of equations, with what is basically Gaussian elimination.
    \end{itemize}
  \end{itemize}
  \subsubsection{Chinese Remainder Theorem}
  Very useful for solving Diophantine equations and systems of congruence equations, which first appeared in ``Sunzi Suanjing'', or ``Mathematical Manual of Sun Zi'', c. 400 AD.
\end{document}
