\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{Lecture 21}
  \subsection{The Medieval Arabic World}
  \subsubsection{The Golden Age}
  Mathematics flourished in the Arabic world during 750--1500 CE.
  \begin{itemize}
    \item They collected, translated, and combined the mathematical knowledge from Greece, India, Egypt, China.
    \item The centralization of resources encouraged people to study math and science, as well as advocacy of ``Hindu-Arabic'' numerals.
    \item New developments in algebra and geometry.
    \item Getting closer and closer to a ``modern treatment'' of mathematics.
    \item Paper became one of the main writing materials, which helped propagate mathematical knowledge.
    \item Around 800, the \emph{House of Wisdom} was founded in Baghdad, and functioned as a knowledge center for many centuries, until it was destroyed in the Siege of Baghdad in 1258.
    \begin{itemize}
      \item It was either a public academy, or private library, or both.
      \item Attracted many mathematicians and contributed to a lot of progress.
      \item A collection and translation of older mathematical texts into Arabic.
    \end{itemize}
    \item Arabic became the standard language throughout the (large) Arabic world, and became the ``unified'' language among scientists (more communication leads to more progress).
  \end{itemize}
  \subsubsection{Number Systems}
  They used the sexagesimal system from the Babylonians, mostly for astronomy, as well as the Hindu-Arabic numeral system which was soon being used for all sorts of mathematics.
  \subsection{Al-Khwarizmi}
  An early director of the House of Wisdom was Muhammad Al-Khwarizmi, who lived from 780--850 CE.
  \begin{itemize}
    \item His name is the origin of our word ``algorithm''.
    \item He strongly advocated for the use of Hindu-Arabic numerals, and it soon became widely used throughout the Arabic world.
    \item Another book he wrote ``The Compendious Book on Calculation by Completion and Balancing'' (830 CE) or ``Al-Jabr'', is the root of our word ``algebra''.
    \item The original ``Al-Jabr'' means ``reunion of broken parts'', ``completion'', or ``rejoining''.
  \end{itemize}
  \subsubsection{Al-Jabr}
  It is the first book known to treat algebra in an elementary way, and for its own sake, and separating it from geometry.
  \begin{itemize}
    \item After being translated to Latin in the 12\tsup{th} century, it became one of the main mathematical textbooks at European universities for over 500 years
    \item The word ``reduction'' means to rewrite an expression in a simplified form, i.e. $x^2 + 4x + 4\to (x + 2)^2$.
    \item The word ``balancing'' means to add or subtract the same quantity to both sides of an equation, and cancel common terms.
  \end{itemize}
\end{document}
