\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{Lecture 6}
  \subsection{Greek Geometry}
  \subsubsection{The Deductive Method}
  The ``Elements'' from Euclid is the first mathematical text where statements are derived step by step, starting from axioms (``postulates'') following certain rules (``common notions'').
  \begin{enumerate}[label=(\arabic*)]
    \item Definitions: It starts with $23$ definitions, i.e. of ``point'', ``line'', ``parallel lines'', ``circle'', etc.
    \item Postulates (``Ruler--Compass Construction''):
    \begin{enumerate}[label=(\alph*)]
      \item To draw a straight line from any point to any point.
      \item To produce a finite straight line continuously into a straight line.
      \item To describe a circle with any center and distance (``radius'').
      \item All right angles are equal to each other.
      \item If a straight line falling on two straight lines makes the interior angles on the same sides less than two right angles, then the two straight lines, if produced indefinitely, meet on the side which the angles are less than two right angles.
    \end{enumerate}
    \item Common Notions:
    \begin{enumerate}[label=(\alph*)]
      \item Things which are equal to the same thing are also equal to each other (``Transitivity of equality'').
      \item If equals are added to equals, the wholes are equal.
      \item If equals are subtracted from equals, the remainders are equal.
      \item Things which coincide with one another are equal to one another.
      \item The whole is greater than the part.
      \begin{note}{}
        The word ``equal'' means ``have the same length/area/volume'', whereas the word ``coincide'' is when two objects are ``really the same''.
      \end{note}
    \end{enumerate}
  \end{enumerate}
  \begin{note}{}
    Often, Euclid uses ``visually obvious'' assumptions that are not mentioned in the postulates.
  \end{note}
  \begin{example}{Proposition I.35 of the ``Elements''} \\
    ``Parallelograms with the same base and the same height are equal''.
  \end{example}
  \begin{note}{}
    Some of the proofs are flawed because they make visual assumptions of the layouts of problems, rather than use equations and variables.
  \end{note}
\end{document}
