% arara: pdflatex
\documentclass[class=article, crop=false]{standalone}
\input{/Users/kylec/github/latex/Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \subsection{Equivalence Relations}
  \begin{definition}{Equivalence relation}
    A relation $R$ on  aset $X$ is an \emph{equivalence relation} if it is reflexive, symmetric, and transitive.
  \end{definition}
  \begin{note}{}
    An equivalence relation gives us a notion of two different elements in a set being ``the same".
    \begin{itemize}
      \item Reflexive: Everything is ``the same" as itself
      \item Symmetric: If $x$ is ``the same" as $y$, then $y$ is ``the same" as $x$
      \item Transitive: If $x$ is ``the same" as $y$, and $y$ is ``the same" as $z$, then $x$ is ``the same" as $z$
    \end{itemize}
  \end{note}
  \begin{example}{Equivalence Relations}
    \begin{enumerate}[label=(\alph*)]
      \item The relation $E$ on the integers where $xEy$ if $x-y$ is even.
      \begin{itemize}
        \item Reflexive: For all $x\in\Z$, $x - x = 0$, which is even, so $xEx$
        \item Symmetric: For all $x, y\in\Z$, if $x - y$ is even, so is $-(x-y)=y-x$. Thus if $xEy$, then $yEx$
        \item Transitive: For all $x,y,z\in\Z$, if $x-y$ is even and $y-z$ is even, then their sum, $x-z$, is also even. Thus if $xEy$ and $yEz$, then $xEz$.
      \end{itemize}
      Observe that this relation relates two integers if they have the same parity.
      \item Let $Y$ be any finite set, and $a,b\in Y^*$ (the set of all strings constructed using $Y$). Consider the relation $L$ over $Y^*$ such that $aLb$ if $a$ and $b$ have the same length.
      \item Let $X$ be the set of all animals, with animals $x,y\in X$. Consider the relation $S$ over $X$ such that $xSy$ if $x$ and $y$ are of the same species.
      \item Let $x,y\in\R$. Consider the relation $C$ over $\R$ such that $xCy$ if $x-y$ is an integer.
    \end{enumerate}
  \end{example}
  \begin{definition}{Equivalence Classes}
    If $R$ is an equivalence relation on a set $X$, then for $x\in X$, the \emph{equivalence class} of $X$ is the set (with respect to $R$), denoted by $[x] = [x]_R = \set{y\in X\mid xRy}$.
  \end{definition}
  \begin{example}{Equivalence Classes}
    \begin{enumerate}[label=(\alph*)]
      \item Let $E$ be a relation on $\Z$, where $xEy$ if $x-y$ is even. The equivalence classes for $E$ are $[0]$ (the evens) and $[1]$ (the odds). So, the set of equivalence classes $= \set{[0], [1]}$.
      \item Let $x, y\in\R$, with the relation $C$ over $\R$ defined by $xCy$ if $x - y$ is an integer. The set of equivalence classes $= \set{[x]\mid x\in[0, 1)}$.
    \end{enumerate}
  \end{example}
  If $R$ is an equivalence relation on a set $X$, then:
  \begin{itemize}
    \item For all $x\in X$, if $x\in [y]$ and $x\in [z]$, then $[y] = [z]$.
    \begin{proof}
      Suppose $x\in [y]$ and $x\in [z]$. Let $w\in [y]$. Because $w\in [y]$, we know that $yRw$. We also know that $yRx$ because $x\in[y]$. By symmetry of $R$, we have $wRy$, and by transitivity, we have $wRx$. But $x\in [z]$, so $zRx$, and by symmetry we have $xRw$. By transitivity, $zRw$ so $w\in [z]$. Thus $[y]\sse[z]$. \par
      By a similar argument, we have that $[z]\sse[y]$, so $[y]=[z]$.
    \end{proof}
    \item For any $x\in X$, $x$ is in some equivalence class, $x\in [x]$ by reflexivity. \\
    So, for every $x\in X$, $x$ is in exactly one equivalence class. If $x$ is in another equivalence class $[y]$, then by the above $[x] = [y]$.
  \end{itemize}
  \begin{definition}{Partition}
    For $X$ a set, a \emph{partition} $\mathcal{S}$ of $X$ is a set of nonempty subsets of $X$ such that every element of $X$ is an element of exactly one of the subsets. In other words, for all $A, B\in \mathcal{S}$
    \begin{itemize}
      \item $A, B\sse X$
      \item $A, B\neq\es$
      \item If $A\cap B \neq\es$ then $A = B$
      \item For all $x\in X$, there exists exactly one $A\in \mathcal{S}$ such that $x\in A$
    \end{itemize}
  \end{definition}
  \begin{note}{}
    We showed that if $R$ is an equivalence relation on $X$ then $\set{[x]_R\mid x\in X}$ is a partition of $X$.
  \end{note}
  \begin{theorem}{Equivalence Relations and Partitions}
    For $X$ a set, there is a bijection $F\colon \text{Set of equivalence relations on $X$}\to \text{Set of partitions of $X$}$, defined by
    \[
      F(E) = \set{[x]_E\mid x\in X},
    \]
    the inverse function $F\inv$ sends a partition $\mathcal{S}$ to the equivalence relation $F\inv(\mathcal{S})$ defined by $xF\inv(\mathcal{S})y$ if and only if $x$ and $y$ are in the same element of $\mathcal{S}$ (in the same equivalence class of $E$).
  \end{theorem}
\end{document}
