\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{Warring States Period}
  \subsection{Warring States and Azuchi-Momoyama Key Dates}
  \begin{itemize}
    \item $1467$--$1477$ \=Onin war
    \item $1542$ Portuguese reach Tanegashima
    \item $1568$ Nobunaga installs last Ashikaga shogun in Ky\=oto (which begins the Azuchi-Momoyama Period)
    \item $1568$ Nobunaga dies
    \item $1573$ Last Ashikaga shogun driven out of Ky\=oto
    \item $1582$ Invasion of Choson (second invasion in 1597)---also known as the Imjin war
    \item $1598$ Hideyoshi dies
    \item $1600$ Battle of Sekigahara
    \item $1603$ Establishments of Tokugawa shogunate
  \end{itemize}
  \subsection{\=Onin War}
  \begin{itemize}
    \item Begins as a conflict over Yoshimasa succession between his son (supported by Yamana clan) and brother (supported by Hosokawa clan)
    \item Shogunate loses all authority
    \item Completely destroys Ky\=oto
    \item Yamana clan leaves the capital in 1477
    \item No clear victor, no one left in charge to guarantee peace between warrior clans
    \begin{note}{}
      The battle started in Ky\=oto, which was supposed to be the one place where there was peace. Once it was destroyed, there was no more centralised authority to keep the peace.
    \end{note}
    \item War sanctions conflicts between warrior domains
    \item No more ``Retired emperors'' (Imperial court impoverished and loses everything)
  \end{itemize}
  \subsection{Socio-economic Developments}
  \begin{itemize}
    \item No effective central authority---many alliances rise up between the numerous warrior clans.
    \item Sh\=oen system has disappeared (in practice but not in name).
    \begin{note}{}
      There are still locations that are called by their estate name, but the estate doesn't really exist any more.
    \end{note}
    \item Increased urbanisation: emergence of villages and towns.
    \item Age of the daimy\=o (warlords).
    \begin{note}{}
      There is economic decentralisation as well---each warlord is receiving taxes from those who live on their land, and doesn't pay taxes to the imperial government.
    \end{note}
    \item Gekokujo: Wealthy peasants become landowners and take samurai status.
    \item Peasant uprisings and village communes.
    \begin{note}{}
      Women no longer own land.
    \end{note}
  \end{itemize}
  \subsection{Socio-military Developments}
  \begin{itemize}
    \item The daimy\=o are completely independent of the government, they issue their own laws for their lands, and form regional alliances.
    \item There are also militarised Buddhist sects:
    \begin{itemize}
      \item Ikk\=o sect (extreme Pure Land sect)
      \item Lotus leagues (followers of Nichiren)
      \item Enyakuji (Tendai sect)
    \end{itemize}
    \item Introduction of firearms in the mid 1500s.
    \item Use of foot soldiers.
    \item Development of castle building---linked to the development of towns.
  \end{itemize}
  \subsection{Buddhist Lineages in the Warring States}
  \begin{itemize}
    \item Most Heian temples are destroyed during the \=Onin war.
    \item Buddhist sects all had armies and fought against domain lords and each other---Lotus v. True Pure Land and Tendai v. Lotus Leagues
  \end{itemize}
  \begin{note}{}
    During the battles, the warriors aren't just fighting for more land---many have the vision of a unified Japan.
  \end{note}
  \subsection{Azuchi-Momoyama Period (1568--1600)}
  \begin{itemize}
    \item Oda Nobunaga
    \item Toyotomi Hideyoshi
    \item Tokugawa Ieyasu
  \end{itemize}
  \subsubsection{Nobunaga}
  \begin{itemize}
    \item $1560$ Wins crucial battle against Imagawa clan (outnumbered $8$ to $1$)
    \item $1561$ Forms alliance with Matsudaira (later Tokugawa)
    \item $1567$ Starts using personal seal: ``Take all under heaven by force''
    \item $1568$ Installs last Ashikaga shogun in Ky\=oto
    \item $1571$ Destroys Enryakuji (Tendai sect)
    \item $1573$ Takeda Shingen dies; Nobunaga drives last Ashikaga shogun out of Ky\=oto
    \item $1575$ Battle of Nagashino (Oda and Tokugawa defeat Takeda)
    \item $1576$ Battle of Tedorigawa (Uesugi defeat Oda)
    \item $1578$ Uesugi Kenshin dies
    \item $1582$ Nobunaga dies at Honn\=oji
  \end{itemize}
  \subsubsection{Hideyoshi (Nobunaga's Successor)}
  Hideyoshi was born a peasant, and rose through the samurai ranks.
  \begin{itemize}
    \item $1585$ Hideyoshi appointed Regent
    \item $1585$--$1587$ Conquers Etch\=u province and Ky\=ush\=u (Shimazu clan)
    \item $1587$ Hideyoshi's edict expelling Christian missionaries
    \item $1588$ Hideyoshi edict prohibiting farmers from bearing arms
    \item $1590$ Conquers H\=oj\=o in Eastern Japan---offers territory to Ieyasu
    \item $1591$ Hideyoshi edicts on census and on restrictions on change of social status
    \item $1592$ Hideyoshi sends army to invade Korea
    \item $1595$ Edict controlling domain lords
    \item $1597$ Second invasion of Korea
    \item $1598$ Hideyoshi dies
  \end{itemize}
  \subsubsection{Tokugawa Ieyasu}
  Son of head of Matsudaira clan (minor warrior clan)
  \begin{itemize}
    \item $1561$ Ieyasu allies with Nobunaga
    \item $1571--1584$ Wars with Takeda clan
    \item $1584$ Ieyasu supports Nobunaga's son Nobuo against Hideyoshi, which ends in a truce
    \item $1590$ Ieyasu accepts Hideyoshi's offer of land exhange
    \item $1598$ Hideyoshi dies
    \item $1600$ Battle of Sekigahara
    \item $1603$ Tokugawa Ieyasu takes title of shogun
  \end{itemize}
\end{document}
