\documentclass[class=article, crop=false]{standalone}
\input{/Users/kylec/github/latex/Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \subsection{Special Functions}
  \begin{definition}{Sequence of elements}
    A sequence in $X$ is a function $s\colon D\to X$ where $D \sse \Z$.
  \end{definition}
  \begin{example}{Sequence}
    \begin{enumerate}[label=(\alph*)]
      \item $X = \set{a, b, c}$, $D = \set{1, 2, 3, 4, 5}$. We may define $s\colon D\to X$ by:
      \begin{align*}
        1 &\mapsto a \\
        2 &\mapsto b \\
        3 &\mapsto c \\
        4 &\mapsto b \\
        5 &\mapsto a
      \end{align*}
      \item The Fibonacci numbers are a sequence of natural numbers. They are defined by: $F_0 = 0, F_1 = 1$, and for $n\geq2$, $F_n = F_{n-1}+F_{n-2}$.
      \item Sequence of even natural numbers: $0, 2, 4, 6, 8, \dotsc$. The function $e\colon \N\to \N$ is defined by $e(n) = 2n$. Observe that the sequence of the powers of 2 is a subsequence of the even natural numbers.
    \end{enumerate}
  \end{example}
  \begin{definition}{Subsequences}
    A \emph{subsequence} of $s\colon D\to X$ is a sequence obtained by restricting the domain of $s$. In other words, a \emph{subsequence} is a sequence of the form $t\colon D'\to X$ where $D' \sse D$.
  \end{definition}
  \begin{definition}{Strings}
    If $X$ is a finite set, a \emph{string} over $X$ is a finite sequence of elements of $X$.
  \end{definition}
  \begin{example}{Strings}
    \begin{enumerate}[label=(\alph*)]
      \item Let $X$ be the English alphabet. Then $c, a, t$ and $d, o, g$ and $m, a, t, h$ are all strings over $X$. We write strings without parentheses and commas, so $c, a, t$ becomes $cat$.
    \end{enumerate}
  \end{example}
  \begin{definition}{Special strings}
    We will let $X^*$ denote the set of strings over $X$. Additionally, let $\lam$ be the null string.
  \end{definition}
  If $\alpha, \beta$ are strings over $X$, we can concatenate them to get a new string $\alpha\beta$.
  \begin{example}{Concatenation}\\
    The string $c, a, t$ concatenated with $d, o, g$ becomes $c, a, t, d, o, g$ or $catdog$.
  \end{example}
  \begin{definition}{Substrings}
    A \emph{substring} is a string obtained by selecting some or all consecutive terms of another string. Observe that the terms must be consecutive, unlike subsequences.
  \end{definition}
  \newpage
  \section{Relations}
  \begin{definition}{Relations}
    A \emph{relation} $R$ from a set $X$ to a set $Y$ is a subset of $X\times Y$. We write $R(x,y)$ or $xRy$ to denote $(x,y)\in R$. If $R$ is a relation from $X$ to $X$, we say that $R$ is a relation on $X$.
  \end{definition}
  \begin{note}{Relations and functions}
    Functions are a special type of relation.
  \end{note}
  \begin{example}{Relations}
    \begin{enumerate}[label=(\alph*)]
      \item Let $X = $ students at UCLA, $Y =$ Classes at UCLA in Winter '21 Quarter. Define $R$ to be a relation between $X$ and $Y$ such that
      \[
      R = \set{(x, y)\in X\times Y\mid x \text{ is a student in }y}.
      \]
      Is $R$ a function? No, because a student can be taking more than one class during the Winter '21 Quarter.
      \item Let $X = \set{2, 3, 4, 5}$ and $Y = \set{4, 5, 6, 7, 8}$. Define the relation $R$ to be: $xRy$ if $x$ divides $y$. Then
      \[
        R = \set{(2, 4), (2, 6), (2, 8), (3, 6), (4, 4), (4, 8), (5, 5)}.
      \]
      \item Let $X = \set{1, 2, 3, 4, 5}$ and define a relation $R$ on $X$ so that $xLy$ if $x < y$. We can visualise this by drawing an arrow $x\rightarrow y$ if $x < y$.
      \begin{center}\begin{tikzpicture}
        \draw(2, 0) circle [radius=0.25] node (A) {$1$};
        \draw[rotate=72](2, 0) circle [radius=0.25] node (B) {$2$};
        \draw[rotate=144](2, 0) circle [radius=0.25] node (C) {$3$};
        \draw[rotate=216](2, 0) circle [radius=0.25] node (D) {$4$};
        \draw[rotate=288](2, 0) circle [radius=0.25] node (E) {$5$};
        \draw[->] (A) -- (B);
        \draw[->] (A) -- (C);
        \draw[->] (A) -- (D);
        \draw[->] (A) -- (E);
        \draw[->] (B) -- (C);
        \draw[->] (B) -- (D);
        \draw[->] (B) -- (E);
        \draw[->] (C) -- (D);
        \draw[->] (C) -- (E);
        \draw[->] (D) -- (E);
      \end{tikzpicture}\end{center}
      \item Let $X = \set{1, 2, 3, 4, 5}$, and define a relation $LE$ on $X$ such that $xLEy$ if $x\leq y$. The diagram is the exact same as above, but every element is also related to itself (because $x\leq x$ for all $x$).
    \end{enumerate}
  \end{example}
  \subsection{Types of Relations}
  \begin{enumerate}[label=(\alph*)]
    \item Reflexive: $R$ is reflexive if for all $x\in X$, $xRx$ ($x$ relates to itself).
    \item Symmetric: $R$ is symmetric if for all $x, y\in X$, $xRy\implies yRx$.
    \item Antisymmetric: $R$ is antisymmetric if for all $x, y\in X$, $xRy$ and $yRx$ implies $x = y$.
    \item Transitive: $R$ is transitive if for all $x, y, z\in X$, $xRy$ and $yRz$ implies $xRz$.
  \end{enumerate}
  \begin{example}{Types of relations}
    \begin{enumerate}[label=(\alph*)]
      \item The relation $<$ over the reals is transitive, (vacuously) antisymmetric, not symmetric, and not reflexive.
      \item The relation $\leq$ over the reals is transitive, antisymmetric, not symmetric, and not reflexive.
      \item Let $X =$ people, and $xNy$ if $x$ and $y$ have the same name. Then $N$ is reflexive, symmetric, and transitive.
      \item Let $X =$ people, and $xTy$ if $x$ is taller than $y$. Then $T$ is transitive, because if $x$ is taller than $y$, and $y$ is taller than $z$, then $x$ is taller than $z$.
    \end{enumerate}
  \end{example}
  \begin{definition}{Inverse of a relation}
    If $R$ is a relation from $X$ to $Y$, then $R\inv$ is the relation from $Y$ to $X$ defined by:
    \[
      R\inv = \set{(y, x)\in Y\times X\mid (x, y)\in R}.
    \]
  \end{definition}
  \begin{definition}{Composition of relations}
    If $R\sse X\times Y$, and $S\sse Y\times Z$, then $S\circ R\sse X\times Z$ such that
    \[
      S\circ R = \set{(x, z)\in X\times Z\mid \text{there exists $y\in Y$ such that $(x,y)\in R$ and $(y,z)\in S$}}.
    \]
  \end{definition}
  \begin{definition}{Equivalence relation}
    A relation is an \emph{equivalence relation} if it is reflexive, symmetric, and transitive.
  \end{definition}
\end{document}
