\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{Lecture 1}
  \subsection{Goals of This Class}
  The end goal of this class is to go over the Fundamental Theorem of Calculus, and to do that we must first cover:
  \begin{itemize}
    \item What is a real number and what makes them special?
    \item How do we define the convergence of a sequence of real numbers?
    \item What is a limit?
    \item What is continuity?
    \item What is a derivative?
    \item What is an integral?
  \end{itemize}
  \subsection{Mathematical Arguments and Logic}
  We have the following components to reasoning about solutions to a mathematical problem:
  \begin{itemize}
    \item Assumptions
    \item Logical steps
    \item Conclusions
  \end{itemize}
  \subsection{Logical Connections}
  We usually use the letters $P$ and $Q$ to denote logical statements (either true or false). We can also use conjunctions to connect logical statements to one another.
  \begin{enumerate}
    \item Conjunctions: ``$P$ and $Q$'', $P\land Q$
    \item Disjunctions: ``$P$ or $Q$'', $P\lor Q$
    \item Implications: ``If $P$, then $Q$'', $P\implies Q$
    \begin{enumerate}
      \item If the proposition is false (i.e. if $P$ is false) then the whole statement is true.
      \begin{definition}{}
        We say that the statement is \emph{vacuously true}.
      \end{definition}
    \end{enumerate}
    \item Negations: ``Not $P$'', $\neg P$
  \end{enumerate}
  \subsubsection{Truth Tables}
  \begin{center}\begin{tabular}{c|c|c}
    $P$ & $Q$ & $P\implies Q$ \\
    \hline
    $T$ & $T$ & $T$ \\
    $T$ & $F$ & $F$ \\
    $F$ & $T$ & $T$ \\
    $F$ & $F$ & $T$
  \end{tabular}\end{center}
  \begin{example}{}
    Prove that if $n$ is an integer, then $n(n + 1)$ is even.
    \begin{proof}
      Suppose that $n$ is an integer. Then we have two cases, where either $n$ is even or $n$ is odd. Let $n$ be an even integer such that $n = 2k$ where $k\in \Z$. Then we have
      \begin{align*}
        n(n + 1) &= 2k(2k + 1) \\
                 &= 2(2k^2 + k).
      \end{align*}
      Thus we see that $n(n + 1)$ is even when $n$ is even. Now let $n$ be odd such that $n = 2m + 1$ where $m\in \Z$. Then we have
      \begin{align*}
        n(n + 1) &= (2m + 1)(2m + 1 + 1) \\
                 &= (2m + 1)(2m + 2) \\
                 &= 2(m + 1)(2m + 1).
      \end{align*}
      Thus $n(n + 1)$ is also even when $n$ is odd, and so is even for all integers $n$.
    \end{proof}
  \end{example}
\end{document}
