\documentclass[class=article, crop=false]{standalone}
\input{../../Preamble}

\fancyhf{}
\lhead{Kyle Chui}
\rhead{Page \thepage}
\pagestyle{fancy}

\begin{document}
  \section{Lecture 2}
  \begin{center}
    \begin{tabular}{c|c|c}
      & Classical Computing & Quantum Computing \\
      \hline
      Software & Boolean Algebra & Linear Algebra \\
      \hline
      Hardware & Classical Mechanics & Quantum Mechanics \\
      & e.g. semiconductors & e.g. superconductors
    \end{tabular}
  \end{center}
  \subsection{The Four Postulates of Quantum Mechanics}
  \subsubsection{State Space Rule}
  The state space is a unit vector with dimension $2^n$, where $n$ is the number of qubits. We use the following abbreviations:
  \[
    \ket{0} = \begin{pmatrix}
      1 \\
      0
    \end{pmatrix}\quad\text{and}\quad \ket{1} = \begin{pmatrix}
      0 \\
      1
    \end{pmatrix}.
  \]
  The state space rule says that we are working with superpositions (or linear combinations).
  \subsubsection{Composition Rule}
  We use the tensor product to compose objects; for each additional qubit the \emph{dimension} of our state space doubles.
  \subsubsection{Step Rule}
  We use unitary matrices to get from one state to the next state. This is because unitary matrices send unit vectors to unit vectors. For example,
  \[
    U\ket{\psi} = \ket{\varphi}.
  \]
  Note that this means that the size of $U$ is $2^n\times 2^n$.
  \subsubsection{Measurement Rule}
  There is a way to get information out of a quantum computer. When we measure a computation performed with $n$ qubits, we get $n$ bits out of the system.
  \subsection{Computations}
  For probabilistic programming, our state is no longer a bit vector, but rather a probability vector. The matrix used to update this state is called a stochastic matrix. We can take the tensor product of two vectors to get the next state.
  \[
    \begin{pmatrix}
      a \\
      b
    \end{pmatrix}\otimes \begin{pmatrix}
      c \\
      d
    \end{pmatrix} = \begin{pmatrix}
      ac \\
      ad \\
      bc \\
      bd
    \end{pmatrix}.
  \]
  \begin{center}
    \begin{tabular}{c|c|c}
      & Probabilistic Computing & Quantum Computing \\
      \hline
      Numbers & Real & Complex \\
      \hline
      State & Vector of Probabilities & Vector of Amplitudes  \\
            & $\sum p_i = 1$ & $\sum \abs{a_i}^2 = 1$ \\
      \hline
      Step & Stochastic Matrix & Unitary Matrix
    \end{tabular}
  \end{center}
  The Hadamard matrix is given by
  \[
    H = \begin{pmatrix}
      \frac{1}{\sqrt{2}} & \frac{1}{\sqrt{2}} \\
      \frac{1}{\sqrt{2}} & -\frac{1}{\sqrt{2}}
    \end{pmatrix}.
  \]
  Observe that
  \begin{align*}
    H\ket{0} &= \frac{1}{\sqrt{2}}\begin{pmatrix}
      1 & 1 \\
      1 & -1
    \end{pmatrix}\begin{pmatrix}
      1 \\
      0
    \end{pmatrix} \\
             &= \begin{pmatrix}
              \frac{1}{\sqrt{2}} \\
              \frac{1}{\sqrt{2}}
             \end{pmatrix} \\
             &= \frac{1}{\sqrt{2}}\ket{0} + \frac{1}{\sqrt{2}}\ket{1}.
  \end{align*}
  We also have
  \begin{align*}
    H\ket{1} &= \frac{1}{\sqrt{2}}\begin{pmatrix}
      1 & 1 \\
      1 & -1
    \end{pmatrix}\begin{pmatrix}
      0 \\
      1
    \end{pmatrix} \\
             &= \begin{pmatrix}
              \frac{1}{\sqrt{2}} \\
              -\frac{1}{\sqrt{2}}
             \end{pmatrix} \\
             &= \frac{1}{\sqrt{2}}\ket{0} - \frac{1}{\sqrt{2}}\ket{1}.
  \end{align*}
  This is analogous to the ``fair flip'' matrix for probabilistic computing. \par
  \subsection{Quantum Computing Trick \#1}
  One of the problems is that many of our operations are non-unitary/irreversible, so how can we use them? With the help of a helper bit $b$, we may define $U_f\colon \set{0, 1}^2\to \set{0, 1}^2$ by
  \[
    U_f(x, b) = (x, b\oplus f(x)),
  \]
  where $\oplus$ is the XOR operator. Via some computation, we find that $(U_f\circ U_f)(x, b) = (x, b)$, so we have successfully made our operation reversible. \par
  This allows us to run the quantum computation for a while. We can then stop it and reverse a few steps to help debug a program.
\end{document}
