\documentclass[class=article, crop=false]{standalone}
\input{/Users/kylec/github/latex/Preamble}

\fancyhf{}
\lhead{Math 61 \\Worksheet 2}
\rhead{Kyle Chui \\Page \thepage}
\pagestyle{fancy}
\pagenumbering{gobble}

\title{Worksheet 2}

\begin{document}
  \maketitle
  \newpage
  \pagenumbering{arabic}
  \begin{problem}{1}
    Let $\set{0, 1}^X$ denote the set of functions $X\to2$ for some set $X$.\par
    Define a function $F$ from $\mathcal{P}(X)$ to $\set{0, 1}^X$ by for $A \in \mathcal{P}(X)$, $F(A)$ is the function $X\to \set{0, 1}^X$ defined by $F(A)(x) = \begin{cases}0 & x\in A \\ 1 & x\notin A\end{cases}$.
    \begin{enumerate}[label=(\alph*)]
      \item List the elements of the set $\set{0, 1}^{\set{a, b}}$.
      \item Let $X = \set{a, b, c}$. Compute $F(\set{b, c})$, $F(\set{a})$, and $F(\set{a, b, c})$. (Remember all the outputs are functions $X\to \set{0, 1}$).
      \item Again let $X = \set{a, b, c}$. Let $g\colon X\to \set{0, 1}$ be the function defined by $g(a) = 1$, $g(b) = 1$, $g(c) = 1$. Find a subset $A$ of $X$ so that $F(A) = g$.
      \item Show that for any set $X$ the function $F\colon \mathcal{P}(X)\to \set{0, 1}^X$ is a bijection (is injective and surjective).
      \item Use this bijection to give another proof that if $X$ is a finite set then $\abs{\mathcal{P}(X)}=2^{\abs{X}}$.
    \end{enumerate}
  \end{problem}
  \begin{enumerate}[label=(\alph*)]
    \item The elements of the set $\set{0, 1}^{\set{a, b}}$ are:
    \[
      \set{(a,0),(b,0)}, \set{(a,1),(b,0)}, \set{(a,0),(b,1)}, \set{(a,1),(b,1)}.
    \]
    \item
    \begin{align*}
      F(\set{b, c}) &= \set{(a, 1), (b, 0), (c, 0)} \\
      F(\set{a}) &= \set{(a, 0), (b, 1), (c, 1)} \\
      F(\set{a, b, c}) &= \set{(a, 0), (b, 0), (c, 0)}
    \end{align*}
    \item Because all elements of $X$ map to 1, we know that none of the elements of $X$ are in $A$. Thus $A = \es$.
    \item
    \begin{proof}
      We will first show that $F$ is injective. Let $A \neq B$. Without loss of generality, there exists some $x\in A$ such that $x \notin B$. Because $F$ is a function, every element in the domain must get mapped to something, so either $(x, 0)$ or $(x, 1)$ is in $F(A)$, but not in $F(B)$. Therefore $F(A) \neq F(B)$ and $F$ is injective. \par
      We will now show that $F$ is surjective. Observe that for every function $f\in \set{0, 1}^{X}$, the set
      \[
        S = \set{x\in X\mid (x, 0)\in f}
      \]
      maps to $f$. Thus $F$ is a bijection.
    \end{proof}
    \item
    \begin{proof}
      Observe that $\abs{\set{0, 1}^X} = 2^X$, because every element in $X$ can map to one of two elements, 0 or 1. We will show that a bijection between finite sets implies that they are of the same cardinality.
    \end{proof}
  \end{enumerate}
  \newpage
  \begin{problem}{2}
    Let $\N_{>1}$ be the set of natural numbers that are bigger than one. For $i \geq 2$, set $X_i = \set{ik\mid k \in \N_{>1}}$. Describe $\N_{>1} \sm \bigpar{\bigcup_{i=2}^\infty X_i}$.
  \end{problem}
  Observe that $X_2$ is the set of all even numbers, $X_3$ the set of all multiples of three, et cetera. Thus the union of all such sets is $\set{x\mid x \text{ is an integer larger than one}}$. Therefore the indicated set is just the empty set.
  \begin{problem}{3}
    Let $f\colon X\to Y$ be a function. Show that $f$ is onto if and only if for every onto function $g\colon Y\to Z$ the function $g\circ f$ is onto.
  \end{problem}
  \begin{proof}
    (\Rightarrow) Suppose $f, g$ are surjective functions. Then for all $z\in Z$, there exists some $y\in Y$ such that $g(y) = z$. Furthermore, for all $y\in Y$ there exists some $x\in X$ such that $f(x) = y$. Thus, for all $z\in Z$, we have $g(f(x)) = (g\circ f)(x) = z$, so $g\circ f$ is surjective. \par
    (\Leftarrow) Suppose $g, g\circ f$ are surjective functions. Then for all $z\in Z$, there exists some $x\in X$ such that $(g\circ f)(x) = z$.
  \end{proof}
\end{document}
